### 2014-12-14 report/algorithms/iterativeLangrange.tex

\begin{algorithm}[H]
\caption{LagrangeIT}
\label{customLagrangeIT}
\begin{algorithmic}
\REQUIRE The matrices $G, Z$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
\ENSURE  Updated $G, Z$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.

\STATE {$s \leftarrow i$}
\STATE {$l \leftarrow j$}

\IF{$g_{ii} > g_{jj}$}
    \STATE {$s \leftarrow j; l \leftarrow i$}
\ENDIF

\STATE{$q \leftarrow \lfloor \frac {g_{ij}}{g_{ss}} \rceil$}

\IF {Verify both $\omega$-Lagrange-reduced conditions}

\STATE{$\vec{z}_l -= q* \vec{z}_s$}
\STATE{$\vec{g}_l -= q* \vec{g}_s$}

\STATE{Updating entries of the Gram matrix}
\ENDIF

\end{algorithmic}
\end{algorithm}