### 2014-12-14 report/figlat/latSVP.tex

\begin{tikzpicture}[scale=0.5]

    \clip (-3,-3) rectangle (5cm,5cm); % Clips the picture...
    \pgftransformcm{1}{0.3}{0.7}{1}{\pgfpoint{0cm}{0cm}}

    \foreach \x in {-14,-12,...,14}{% Two indices running over each
      \foreach \y in {-14,-12,...,14}{% node on the grid we have drawn 
        \node[draw,circle,inner sep=1.2pt,fill=mygreen] at (\x,\y) {};
            % Places a dot at those points
      }
    }
  \node (b) at (0.5,0) {$\cal O$} ;

\node[draw,circle,inner sep=17pt,fill=myblue, opacity =.30] at (0,0) {};

  \end{tikzpicture}