### 2014-12-19 report/paper.tex

\documentclass[10pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{color}
\usepackage{graphicx}
\newcommand{\my}[1]{{\color{blue} #1 }}
\setlength{\parindent}{0pt} 
\floatname{algorithm}{Procedure}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\usepackage{tikz}
\usetikzlibrary{shapes,decorations.pathreplacing,calc,shapes.callouts,decorations.pathmorphing}

\usepackage[normalem]{ulem}
\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}
\definecolor{mygreen}{rgb}{.222,.62,.49}

\title{Comparing the Jacobi Method and LLL lattice reduction algorithm\my{s} for cryptographic applications}
\date{Fall 2014}
\author{Frederic Jacobs\\ EPFL Bachelor Semester Project\\ frederic.jacobs@epfl.ch}

\renewcommand{\vec}[1]{\mathbf{#1}}

\begin{document}

\maketitle

\section{Introduction}
In 2012, Sanzheng Qiao published a paper~\cite{originalJacobiMethodLatticeBasisReduction} proposing the use of the Jacobi method for lattice reduction applications. In following papers, Qiao proposed several enhancements to the generic Jacobi method. \my{The purpose of their research is to generate a reduction that produces good reduction on random matrices for which the columns represent almost orthogonal vectors. This is in contrast to lattices that are used in cryptography as public keys. From the cryptanalytic point of view, we are interested in random lattices where the given basis vectors are skewed and not almost orthogonal.} We implemented the Jacobi method reduction algorithm \my{as presented by Qiao et al. and compared the quality of the outputs with the LLL algorithm for almost orthogonal matrices and random lattices. The results show...}

\section{Preliminaries}

In this section, we briefly cover the basics of lattices and lattice reduction required in the scope of this report.

\subsection{Euclidean norm}

The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n $ \my{to} $\mathbb{R}$ that assigns \my{a} length to any vector of the space \my{$\mathbb{R}^n$}.

The mapping is defined for any vector $\vec{b}$ of $\mathbb{R}^n$ with coordinates $b_i$ as follows:

\[
\|\mathbf{b}\| = \sqrt{b_1^2 + b_2^2 + ... + b_n^2} 
\]

which results from the Pythagorean theorem.

\subsection{Lattice, also called Euclidean lattice}

Informally, a lattice can be defined as the infinite set of intersection points of an infinite regular grid as shown in Fig.~\ref{fig:latti}.
\begin{figure}
\centering

\input{figlat/latticegrid}\hspace{2cm}
\input{figlat/latticeZ}

\caption{\my{text text text}}
\label{fig:latti}
\end{figure}

This can be translated into a more formal definition~\cite{SchnorrStanfordNotes} \my{as follows}.\newline 
Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the discrete additive subgroup

\[
L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
\] of $\mathbb{R}^m$ a lattice with basis $\vec{b_1,b_2,...,b_n}$.
A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors is fully described by this basis which we denote in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.

A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$. We will work exclusively with full dimensional lattices. This means that we will work on matrices with $m=n$ linearly independent vectors \my{in $\mathbb{R}^n$}.

\subsubsection{Equivalent bases}

As this article focuses on lattice basis reduction, it is important to define the equivalence of two bases. Two bases $B_1 , B_2$ are equivalent iff $B_2 = B_1 U$ for a unimodular matrix $U$.
A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$. \my{$B_1$ and $B_2$ are bases to the same lattice.} The right multiplication of a matrix $B_1$ with a unimodular matrix $U$ yields $B_2$ by performing a combination of the following \my{elementary integer} column operations:
\begin{itemize}
\item Swap two columns
\item Negate the elements of a column
\item Addition of an integer multiple of one column to another
\end{itemize}

A lattice can be described by infinitely many equivalent bases. The dimension, the length of a shortest vector and the volume, as described next, stay invariant under change of basis.
\subsubsection{Volume}

The volume of a lattice, also known as the determinant of the lattice, is defined as:
\[
\text{vol} \,L := \sqrt{\det(B^T B)}
\] where $B$ is a matrix of the lattice's basis vectors.

The determinant of an unimodular matrix being $\pm 1$, the determinant in the above equation is an invariant under change of basis. It follows that the volume of a lattice, vol$(L)$, is an invariant under change of basis.

Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant of the basis matrix.
For full dimensional lattices, the volume can be computed:
\[
\text{vol} \,L_{\text{full}} = |\det B|
\] where $B$ is a matrix of the linearly independent lattice basis vectors.

\subsubsection{Hermite Normal Form}

The \emph{Hermite Normal Form}~\cite{SchnorrStanfordNotes} representation of a basis matrix $B$, abbreviated \emph{HNF(B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.

A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when:
\begin{enumerate}
\item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
\item $a_{ii} > 0$ for $i=1,2,...,n$
\item $0 \leq a_{ij} < a_{ii}$ for $j < i$
\end{enumerate}

\subsection{Two types of lattices}
For our experiments, we will take into account two types of lattices: Random lattices as used in cryptology and special lattices that allow an almost orthogonal basis.

\subsubsection{Random Lattice}

We say that a lattice is a random lattice $L$ of prime volume $P$ if under HNF form its basis matrix $B$ has the following properties:

\begin{itemize}
\item the diagonal has 1 for all its entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
\item All row entries of the matrix right to the position that is set to $P$ are smaller than $P$ in absolute value.
\end{itemize}

Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows: \my{no bold ai's in the matrix}
$$\left[\begin{array}{ccccc}
P & \vec{a_2} & \dots & \vec{a_{m}}\\
& 1&   & \\ 
& & \ddots & \\ 
 & & &1 
\end{array}\right]$$
where $|a_i| < P$.

\subsubsection{Almost Orthogonal Lattice Bases}

We define an \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as an $n \times n$ square matrix whose entries are $k$-bit integers picked at random. These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.

\subsection{Lattice Reduction}

Lattice reduction is the process of finding nearly orthogonal vectors from an integer lattice basis. Giving an exact definition of "nearly orthogonal vectors" is tricky and each reduction algorithm is likely to give it's own definition of what it means for a basis to be reduced.

\subsection{Quality of a reduction}

In this section, we will discuss two lattice reduction quality indicators that we will use to compare the result of reduction algorithms.

\subsubsection{Orthogonality defect of a basis}
\label{sec:orthogonalityDefect}
The \emph{orthogonality defect} of a basis $\vec{b_1},\vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by:
\[
    \text{OrthDefect}(L) :=  \sqrt[n]{\frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}}
\]

The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}. As the basis vectors are getting \my{increasingly? more} orthogonal during a reduction, the nominator will become smaller. The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$. In the special case where the basis would be completely orthogonal, the orthogonality defect would be $\text{OrthDefect} = 1$.

\subsubsection{Hermite factor of a basis}
\label{sec:hermiteFactor}
The \emph{Hermite Factor} of basis vectors $\vec{b_1}, \vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by

\[
    \text{HF}(L) := \frac{\|\vec{b_1}\|}{\sqrt[n]{\det(L)}}
\]

The Hermite Factor is a quality indicator that we will use to compare the reduction to a \emph{shortest vector}.

\subsection{Lattice Hard Problems}

Lattices have a few known hard problems that can be used for asymmetric cryptography. We will only cover the most famous one, the Shortest Vector Problem.

\subsubsection{Shortest Vector Problem (SVP)}

Given a basis of a\my{n} $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$

In other words, we are looking for the shortest vector of the lattice. Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions~\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.

\subsubsection{Minkowski's Convex Body Theorem}
\my{Still check grammar in following sentence.}
For a full-rank lattice $L$ of $\mathbb{R}^n$ and $C$, a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$. Then $C$ contains at least a nonzero point of the lattice $L$.

A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero lattice-vector $x$ whose length is bounded by $\sqrt{n} (\det L)^{1/n}$. So we can derive a bound on the length of a shortest lattice-vector:
\[
\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}.
\]
\my{In the 80's lattice reduction showed to be successful in finding shortest vectors whose length was significantly smaller than indicated by the above bound.}

\subsection{Gram-Schmidt orthogonalization process}
The \emph{Gram-Schmidt orthogonalization (\emph GSO) process} is a method for orthonormalising a set of vectors. Two vectors are orthonormal if they are both orthogonal and unit vectors.

The GSO process  can be defined recursively:
\[
\begin{cases}
\vec{b_1}^{\star} = \vec{b_i}  \\
\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b_j^\star}
\end{cases}
\]
where $\mu_{ij} = \frac{\langle \vec{b}_i, \vec{b}_j^\star \rangle}{\| b_j^\star\|^2}$. The set of vectors $b_i^*$ are only constituting a basis of the lattice if their $b_i^* \in \mathbb{Z}$ \my{. No, if $\mu $'s are integers.}

\section{The LLL Algorithm}

The LLL algorithm~\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis. LLL and BKZ 2.0 are the two reduction algorithms that are used in practice. LLL is used in a first step and BKZ, with a suitable blocksize, is then used in a second step to reduce the basis further if needed.

\subsection{$\delta$-LLL Reduced}
We say that a basis is $\delta$-LLL reduced if:
\begin{itemize}
\item $| \mu_{i,j}| \leq \frac{1}{2} $ for $1 \leq j < i \leq n$
\item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
\end{itemize}

The first condition guarantees the length reduction of the ordered basis.
The Lovász conditions assures us that although the Gram-Schmidt vectors can get shorter and shorter, their length in a sorted basis cannot decrease too quickly.

Since the LLL-reduction algorithm returns a $\delta$-LLL reduced basis, we can observe the following properties on the basis:
\begin{itemize}
\item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
\item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$ \my{det(B) or det(L) ?  }
\end{itemize} 
where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha > 4/3$.

\subsection{LLL overview}

The LLL algorithm, like many reduction algorithms, can be broken down into two parts: size-reduction and vector swapping. Size-reduction reduces the vectors from the input basis. After that, the vectors of the basis are checked against the Lovász condition. If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't \my{a pair is singular} satisfy the condition, we swap \my{the vectors} and go back to the size reduction (cfr. \textbf{Procedure \ref{alg:lllAlgorithm}}).

\begin{algorithm}[H]
\caption{Overview of naive-iterative LLL}
\label{alg:lllAlgorithm}
\begin{algorithmic}[1]
\REQUIRE a basis ($\vec{b_1, ..., b_n}$) of a lattice $L$.
\ENSURE $\delta$-LLL reduced basis $(\vec{b_1, ..., b_n})$

\STATE Size-reduce $(\vec{b_1, ..., b_n})$
\IF{there exists an index $j$ which does not satisfy Lovász' condition} \STATE{swap $\vec{b_j}$ and $\vec{b_{j+1}}$, go back to step 1} \ENDIF

\end{algorithmic}
\end{algorithm}

\section{The Jacobi Method}

In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction~\cite{originalJacobiMethodLatticeBasisReduction}. The idea of the Jacobi method is to reduce larger dimension lattice bases by using the Lagrange algorithm iteratively.

\subsection{Lagrange-reduced basis}

Lagrange was the first to formalize the notion of reduction for rank-two lattices~\cite{lagrangeArithmetique}.

Let $L$ be a rank-two lattice of $\mathbb{R}^n$. A basis $(\vec{b_1}, \vec{b_2})$ of $L$ is said to be Lagrange-reduced if and only if $\|\vec{b_1}\| \leq \|\vec{b_2}\|$ and $| \langle \vec{b_1} \vec{b_2} \rangle | \leq \| \vec{b_1} \|^2 / 2 $.

\subsection{Lagrange algorithm}
\label{sec:lagrangeAlgorithm}

Lagrange's algorithm solves the shortest vector problem exactly for lattice bases of dimension 2. It is a variant of Euler's centered algorithm that is used to compute the greatest common divider \my{still not divisor?} of two numbers. GCD can be thought to be a one dimension reduction problem. The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.

\input{algorithms/lagrange}
\subsection{The Generic Jacobi Method}
We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice. This method was introducedby Qiao in 2012~\cite{originalJacobiMethodLatticeBasisReduction} claiming experimental results to be considerably faster than LLL \my{for almost orthogonal bases}.

From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis for which:
\begin{align}
\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
\end{align}

\input{algorithms/genericJacobi}

\subsection{A Jacobi Variant}

In this part, we present the Jacobi-variant that was used for running our benchmarking.

\subsubsection{$\omega$-Lagrange-reduced conditions}
There are two conditions for a basis to be $\omega$-Lagrange-reduced expressed with the Gram matrix $G$ of the basis $B$. The Gram matrix $G$ of a basis $B$ is computed \my{as} $G = B^T B$. By construction, a Gram matrix is always symmetric, for every pair of indices $(i,j)$, we have $g_{ij} = g_{ji}$. \my{A basis is $\omega$-Lagrange-reduced if it satifies: }

\[
\begin{cases}
\lfloor g_{sl} / g_{ss} \rceil \leq 1, \\
\omega^2 g_{ll} \leq g_{ss} + g_{ll} - 2|g_{sl}|

\end{cases}
\]  where $1/\sqrt{3} \leq \omega < 1$ \my{and $1\leq s,\ell \leq n $}. The $s$ and $l$ indices are referring to the index of the short and long vectors in the basis Gram matrix of the basis.

\subsubsection{Size-reduction Lagrange}
The size-reduction algorithm \my{better LagrangeIT} is a variant of the Lagrange algorithm that we presented in section~\ref{sec:lagrangeAlgorithm}. The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied. In which case, both the unimodular reduction matrix $Z$ and the Gram matrix $G$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.

\input{algorithms/lagrangeIT-full}

\subsection{Jacobi-variant algorithm}
This Jacobi-variant performs Lagrange reductions iteratively on the Gram matrix until no more reduction is done. That is when both $\omega$-reduced Lagrange conditions are satisfied.

\input{algorithms/fast-Jacobi}

\section{Experimental Results}

\subsection{Our Implementation}

We implemented the Fast-Jacobi\my{~\cite{}} Jacobi reduction\my{, Proc.\ref{jacobiVariantReduction},} in C++ on top of the \emph{newNTL} library. The FPLLL implementation of LLL was used to compare the results after reduction.

\subsection{Scope of testing}

Our experimentation focused on the quality of the reduction. The two quality \my{metics?} we used for comparing the reductions are:

\begin{itemize}
\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice \my{in comparison to the volume}.
\item the Hermite factor (See section: \ref{sec:hermiteFactor}), gives us an indication about the size of the shortest vector of the reduced basis for a given lattice.
\end{itemize}

\subsection{Input bases}
Numerous papers were written about the Jacobi method for lattice reduction by S. Qiao , namely~\cite{originalJacobiMethodLatticeBasisReduction}, \cite{complexityAnalysisOfJacobiMethod} and ~\cite{enhancedJacobi}, but none of the papers \my{is} clarifying what bitsize was used for the randomly generated entries of the vectors. The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance. MIMO uses smaller dimensions lattices. In \cite{enhancedJacobi} and \cite{tian2014hybrid}, a reduction is performed on lattices of dimensions 20 to 100. These dimensions are inappropriate for cryptographic applications.

In our experiments, we have used a bitsize of around 10 bits entries.

\subsection{On Almost Orthogonal Basis \my{plural?}}

We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results. The orthogonality defects are similar to what the authors found in \cite{tian2014hybrid} and \my{\cite{fastJacobi}}. Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}. By increasing the parameter $\omega$, we are making the $\omega$-Lagrange-reduced conditions more restrictive. More specifically, for every pair of vectors of the basis, the following condition needs to be satisfied
\[
 \omega \| \vec{a}_l \| \leq \|\vec{a}_l \pm \vec{a}_s \|
\] where $\vec{a}_l$ is the longest vector and $\vec{a}_s$ is the shortest. Hence, the reduction renders the basis more orthogonal as $\omega$ grows. \my{We see that for $\omega=0.6$ , LLL outperforms... while for $\omega = 0.9$...}

\begin{figure}[H]
\begin{minipage}{.45\textwidth}
  \centering

    \includegraphics[width=0.9\textwidth]{results-graphs/random-matrix-omega-06defect.png}

      \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
  \label{fig1}
\end{minipage}
\hspace{.05\textwidth}
\begin{minipage}{.45\textwidth}

  \centering

    \includegraphics[width=0.9\textwidth]{results-graphs/random-matrix-omega-09defect.png}

  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
  \label{fig2}
\end{minipage}
\end{figure}

The increase in the parameter $\omega$ comes at a significant performance cost as illustrated in Fig~\ref{fig3}.

\begin{figure}[H]
\centering
\includegraphics[width=0.5\textwidth]{results-graphs/runningtime.png}

\caption{Running time increasing substantially with $\omega$ parameter}
\label{fig3}
\end{figure}

\subsection{On Random Lattices}

Our experiments on random lattices on researched dimensions ($> 50$) have been aborted due to excessive running times.

\subsection{Jacobi after LLL}

As we found that Jacobi could reduce a random lattice further than LLL when $\omega$ is close to 1, we tried to run the Jacobi reduction algorithm after LLL reduction of a random lattice to see if it was able to reduce it further. Experimental results (Fig.~\ref{fig4}-\ref{fig5}) are showing that the Jacobi method does indeed reduce the bases further.

\begin{figure}[H]
\begin{minipage}{.45\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{results-graphs/LLLThenJacobi/defect.png}
      \caption{Jacobi after LLL}
  \label{fig4}
\end{minipage}
\hspace{.05\textwidth}
\begin{minipage}{.45\textwidth}

  \centering
    \includegraphics[width=0.9\textwidth]{results-graphs/LLLThenJacobi/defect-logscale-difference.png}

  \caption{Logarithmic-scale difference in orthogonality defects}
  \label{fig5}
\end{minipage}
\end{figure}

We notice that the Jacobi algorithm does reduce the bases further than LLL. Looking at the conditions to be $\delta$-LLL reduced and $\omega-$Jacobi reduced, we can find for example two bases that are LLL reduced but not Jacobi reduced \my{and will miss the shortest vector}.

\begin{enumerate}

\item \[
B = \begin{bmatrix}
  \vec{b_1} \\
  \vec{b_2} \\
  \vec{b_3}
 \end{bmatrix} = \begin{bmatrix}
  0 & 2 & 0 \\
  0 & 1 & 2 \\
  2 & 0 & 0
 \end{bmatrix}
\]

\item \[
B = \begin{bmatrix}
  \vec{b_1} \\
  \vec{b_2} \\
  \vec{b_3}
 \end{bmatrix} = \begin{bmatrix}
  4/3 & 0 & 0 \\
  2/3 & \sqrt{4/3} & 0 \\
  2/3 & \sqrt{4/3}/2 & 1
 \end{bmatrix}
\]

In the first case, LLL will stop while Jacobi does exchange $\vec{b}_2$ and $\vec{b}_3$. In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
\end{enumerate}

\subsection{Note on running times}
Although the analysis of the running time was not a metric that was researched in this paper, we did notice that the Jacobi method running time was growing substantially with the dimensions. We were unable to reproduce reduction running times lower than those of LLL as reported by the authors.

\section{Further research}
As follow-up of this research, we believe it would be useful to:
\begin{itemize}
\item determine the algorithmic complexity of the Jacobi method
\item benchmark the Jacobi method against BKZ after LLL
\end{itemize}

\section{Conclusions}

Although the Jacobi method might have practical applications in signal processing, it is not appropriate for cryptographic applications on random lattices or higher dimension\my{al} bases.

\section{Acknowledgments}
I would like to particularly thank Anja Becker and Nicolas Gama from LACAL who have been providing significant feedback and advice during my research.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}